### 2020-12-22 latex/acknowledgements.tex
Final edit, she be complete!

    These two years of doing research together were the most fun I've ever
    had; yet they've also been fraught with death, sorrow, and the most
    devastating heartbreak I've ever experienced.
    Through it all you've been nothing but kind, honest, generous,
    compassionate, understanding, and so incredibly humble.
    You work unbelievably hard in the background to ensure your students succeed, and you
    ask for so little in return.
    My name is on the title page of this thesis, but if not for the required
    academic formality I would change the authorship to read
    ``Adan Moran-MacDonald, who couldn't have done any of this without the love
    and support of his supervisor and dear friend, Manfredi Maggiore."

    Manfredi: you are my hero.

    % TODO: do the rest of the acknowledgements: parents, SCG friends, other
    % friends, committee

>>>>

    These two years of research together were the most fun I've ever
    had; they were also fraught with death, sorrow, and the most
    devastating heartbreak I've ever experienced.
    Through it all you've been nothing but kind, honest, generous,
    compassionate, understanding, and so incredibly humble.
    You work tirelessly in the background to ensure your students
    succeed, yet ask for so little in return.
    My name is on the title page of this thesis, but if not for the required
    academic formality I would change the authorship to read
    ``Adan Moran-MacDonald, who couldn't have done any of this without the love
    and support of his supervisor and dear friend, Manfredi Maggiore."

    Manfredi: you're my hero.

    \bigskip % big line break to separate

    My gratitude also goes to Lacra Pavel, Raymond Kwong, and Stark Draper for
    being on my committee. I hope you enjoy reading this thesis. Thank you for
    your hard work.

    Dear friends in the Systems Control Group,
    you remind me what it means to have a zest for life. 
    Every moment with you is wild, crazy, never-ending fun. Thank you
    for the memes, deep chats, pub games, and bean juice; the distractions
    always came when I needed them most.

    Oh family of mine, your unconditional support has shaped every word I've
    written in this thesis, and you deserve more appreciation than I can put
    into words. Thank you for driving me to be myself. I love you all.

    Finally, to Mallory, Leela, and Fry: our life together may have ended, but
    you filled my heart and helped me become who I am today. 
    Thank you for the happiest years of my life.

### 2020-12-22 latex/acrobot.tex
Final edit, she be complete!

Defining \(M(q) := D_s(q)\) and \(V(q) := P_s(q)\), we find the conjugate of momenta 
is \(p = (p_u, p_a) = M(q)\dot{q}\).
The dynamics in \((q,p)\) coordinates are given by
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p \\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) \\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
where the inverse inertia matrix is
\begin{equation}\label{eqn:Minv}
    \Minv(q) = \frac{1}{ml^2\left(2-c_a^2\right)}
    \begin{bmatrix}
        1 &
        -\left(1+c_a\right) \\
        -\left(1+c_a\right) &
        3+2c_a
    \end{bmatrix}
    .
\end{equation}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates inside the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
:= \Sone \times \Sone\), and
\(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
:= \R \times \R\).
This allows us to apply the theory of VNHCs from Chapter
\ref{ch:vnhcs}.

@@

In the spirit of perturbation analysis, we will expand our time-scaled system
\(\hat{\mu}(\alpha,\mu_0,I)\).
We begin by setting \(I = 0\) to find the nominal system
\[
\begin{cases}
    \pdiff{\hat{\mu}_0}{\alpha} = g(\alpha,\hat{\mu}_0,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
   .
\end{cases}
\]
Equations ~\eqref{eqn:acrobot-mu-dot-nom}-\eqref{eqn:acrobot-alpha-dot-nom}
reveal that \(g(\alpha,\mu,0) = 0\),
so the solution to this nominal system is 
\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies
\[
    P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma
    ,
\] 
when \(I < 0\). 
In this case, the constraint dissipates energy on \(\mathcal{O}_1\).

@@

Indeed, let \(E_\pi\) be the level set with energy \(E(\pi,0)\), 
which forms the boundary of \(\mathcal{O}_1\). 
Define \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) to be the closure of
\(\mathcal{O}_1\).
An orbit of the acrobot will only begin rotating if it escapes 
\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\).
It is possible for orbits starting in \(\mathcal{O}_1\) to always approach
\(E_\pi\) without ever crossing into the rotation zone.
We will prove this does not happen in general.

@@

Escaping once into the rotation domain does not guarantee the acrobot
orbit will not return to \(\mathcal{O}_1\).
Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
several times.
However, the orbit must eventually remain outside the closure of 
\(\mathcal{O}_1\). 
If not, the positive limit set would once again be
\(E_\pi\); by the same argument as above, the orbit must converge to 
\((\pi,0)\).

@@

Note that the boundary of \(\mathcal{R}\) intersects the \(p_u\) axis when 
\(p_u^2 = 60m^2 g l^3\). 
This boundary is the homoclinic orbit for the
upright equilibrium of the pendulum, which is the level set with energy
\(E(\pi,0)\).
Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\), if a
rotation has momentum \(p_u > 0\), 
and \(\rho < -\sqrt{60 m^2 g l^3}\) if it has momentum \(p_u < 0\).
The energy level set associated with \(\rho\) is 
\[
    \left\{(q_u,p_u) \in \SxR \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:rotation-pu2}
    \frac{p_u^2}{10m l^2} + 30mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
    .
\end{equation}
On this level set, \(q_u\) takes all values on \(\Sone\), so our angle of
rotation is uniquely parameterized by \(\beta = q_u\).
Since \(\rho\) does not change sign along the rotation, we have the smooth
relationship
\begin{align}\label{eqn:rotation-Tinv}
    q_u &= \beta
    , \\
    p_u &= \sign{\rho}\sqrt{\rho^2 - 30 m^2 g l^3\left(1 - c_\beta \right)}
    .
\end{align}

@@

Computing the acrobot's constrained dynamics in \((\beta,\rho)\)-coordinates and
setting \(I = 0\) yields the dynamics of the nominal pendulum.
MATLAB evaluates those dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\rho} &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\beta} &=  \sign{\rho} 
    \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
    .
\end{align}
As expected, \(\dot{\beta}\) does not change sign on \(\mathcal{R}\) because the
orbits always flow clockwise.
If \(\rho > 0\), the rotation curve goes from \(\beta = -\pi\) to 
\(\beta = \pi\); 
if \(\rho < 0\), it goes from \(\beta = \pi\) to \(\beta = -\pi\).
By continuity of ~\eqref{eqn:rotation-T}, there exists \(I_1 > 0\) small enough
that \(\dot{\beta}\) is non-zero and does not change sign on \(\mathcal{R}\) for
all \(I \in [-I_1,I_1]\).

@@

Let us pause here for a moment to remember what we are trying to accomplish.
The second part of Theorem \ref{thm:acrobot-energy-stabilization} states that,
under suitable conditions on the integral of \(b(\beta,\rho_0)\), 
the acrobot will gain energy on 
\[
    \mathcal{O}_2(\bar{\rho}) := \left\{ (q_u,p_u) \in \SxR
    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,
\]
for some \(\bar{\rho} > \sqrt{60m^2gl^3}\).
Recall from Section \ref{sec:acrobot-proof-o1} that any acrobot initialized in
\(\mathcal{O}_1\) will eventually enter the rotation region \(\mathcal{R}\).
What remains, then, is to prove that the acrobot's orbits will also gain
energy on
\[
    \mathcal{R}_{\bar{\rho}} := \mathcal{R} \cap \mathcal{O}_2(\bar{\rho})
    .
\]
This can be proven by analyzing the Poincar\'{e} map \(P_\mathcal{R}(\rho_0)\).
Looking at Figure \ref{fig:acrobot-rhobar-regions}, it is clear that
\(P_\mathcal{R}\) has the domain
\[
    D := \left[-\bar{\rho},\sqrt{60m^2gl^3}\right[ \, 
    \cup \, 
    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
which is the intersection of \(\mathcal{R}_{\bar{\rho}}\) with the \(p_u\)-axis.
We label the part of \(D\) contained in the negative \(p_u\)-axis as
\[
    D^- :=\left[-\bar{\rho},\sqrt{60m^2gl^3}\right[
    ,
\]
and the part contained in the positive \(p_u\)-axis as
\[
    D^+ := \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
so that \(D = D^- \cup D^+\).
If \(P_\mathcal{R}(\rho_0)\) is always further from the origin than 
\(\rho_0\) (\ie~ it is ``expanding" on \(D\)), the acrobot will be gaining energy
on \(\mathcal{O}_2(\bar{\rho})\).

@@

\subsubsection*{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
\(\hat{\rho}(\beta,\rho_0,I)\) by studying the evolution of the discrete time
ODE
\begin{equation}\label{eqn:rhohat-discrete}
    \rho_{n+1} := P_\mathcal{R}(\rho_n) 
    = \rho_n + I\sign{\rho_n}S(\rho_n) + 
        R(\sign{\rho_n}2\pi,\rho_n,I)
    .
\end{equation}
Here, \(\rho_n\) represents the distance along the
\(p_u\)-axis when an orbit of the constrained dynamics intersects the
\(p_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
\((q_u,p_u) = (0,\rho_0)\).

@@

We begin with \(\rho_0 \in D^+\).
Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
its parameters.
Therefore, it is bounded below on the compact set
\(\rho_0 \in \left[\sqrt{60m^2gl^3},\bar{\rho}\right]\) by some value
\(\underbar{R}(I)\).
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_2 > 0\) and \(r > 0\) so that, for all \(I \in [-I_2,I_2]\),
\[
    R(2\pi,\rho_0,I) \geq \underbar{R}(I) > -I^2r
    .
\]
Since \(S(\rho_0) \geq \epsilon\), we find that
\[
    P_\mathcal{R}(\rho_0) > \rho_0 + I\epsilon -I^2r
    ,
\]
for all \(\rho_0 \in D^+\).
Choosing \(I_3 \in \,]0,I_2]\) so that
\[
    I_3\epsilon - (I_3)^2 r \geq \gamma
    ,
\] 
makes \(P_\mathcal{R}(\rho_0) \geq \rho_0 + \gamma\) for all 
\(\rho_0 \in D^+\)

Now take \(\rho_0 \in D^-\), where we wish to show that 
\(P_\mathcal{R}(\rho_0) \leq \rho_0 - \gamma\).
The remainder \(R(\beta,\rho_0,I)\) is bounded above on the compact set 
\(\rho_0 \in \left[-\bar{\rho},-\sqrt{60m^2gl^3}\right]\), and 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_4 > 0\) and \(r^\prime > 0\) so that, for all \(I \in [-I_4,I_4]\), 
\[
    R(-2\pi,\rho_0,I) < I^2 r^\prime
    .
\]
Hence,
\begin{align*}
    P_\mathcal{R}(\rho_0)
    &= \rho_0 - IS(\rho_0) + R(-2\pi,\rho_0,I)
    , \\
    &\leq \rho_0 - I\epsilon + I^2r^\prime
    .
\end{align*}
Choosing \(I_5 \in ]0,I_4]\) so that 
\[
    I_5 \epsilon - (I_5)^2r^\prime \geq \gamma
    ,
\] 
makes \(P_\mathcal{R}(\rho_0) \leq \rho_0 - \gamma\) for all 
\(\rho_0 \in D^-\).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies 
\[
    \abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma
    ,
\]
when \(I < 0\).

@@

\section{Experimental Results}
In this section, we test our VNHC on the physical acrobot built by
\citet{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
Since it is not a simple acrobot, this experiment will test whether VNHCs are
robust to model uncertainties.
As we will see, our VNHC can inject energy on the general acrobot model.

@@

\begin{table}
    \centering
    \caption{Physical parameters for the real acrobot, as measured by
    \citet{xingbo_thesis}.}
    \label{tab:acrobot-parameters}
    \begin{tabular}{ccccccccc}
        \toprule
        $m_t$ & $m_l$ & $d_t$ & $d_l$ & $l_t$ & $l_l$ & $J_t$ & $J_l$ & $g$ \\
        \midrule
        0.2112 & 0.1979 & % mass
        0.148 & 0.145 & % total length
        0.073 & 0.083 & % COM distance
        0.00129 & 0.00075 & % moment of inertia
        9.81 \\ % gravity
        \bottomrule
    \end{tabular}
\end{table}

@@

To compute the energy of the nominal pendulum, we evaluate the system at the
VNHC \(q_a = 0\), which yields
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Hence, the level set of \(E_\pi\) of the energy \(E(\pi,0)\) is parameterized by
\begin{align*}
    p_u &\approx \pm \sqrt{\frac{1.1994}{793.1001}(2 - (1 - \cos(q_u))}
    ,
\end{align*}
for each \(q_u \in ]-\pi,\pi]\).

@@

\subsection{Physical Experiments}
To complete this chapter, we perform a four experiments with the physical
acrobot:
\begin{enumerate}
    \item An undisturbed test to validate that the acrobot reaches rotations.
    \item A test where we stop the acrobot once it has started rotating.
    \item A test where we push the acrobot in its direction of motion after it has
        started rotating.
    \item A test where we push the acrobot against its direction of motion.
\end{enumerate}

@@

\subsubsection*{Test 2: Stopping the Acrobot}
For this test we initialize the acrobot at 
\((q_u,p_u) \approx \left(\pi,0\right)\) and let it run for 15
seconds, then stop the acrobot as it reaches the bottom of its arc.
The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
The blue curves correspond to the orbit before the disturbance, while the red
spiral shows that the acrobot continues to oscillate after it is stopped.
Despite the disturbance, it gains energy and eventually reaches a rotation
again.

@@

\subsubsection*{Test 3: Pushing the Acrobot Forwards}
To see how the acrobot responds when pushed in its direction of motion, we allow
the acrobot to rotate freely for 15 seconds and then give it a push in its
direction of motion.
The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot speeds
up to rotate with energy \(E(0,0.22)\), but then slows down until it reaches a
stable rotation with energy \(E(0,-0.195)\).

@@

\subsubsection*{Test 4: Pushing the Acrobot Backwards}
In this final experiment, we test whether the acrobot can easily change
directions when pushed against its current direction of motion.
We again allow the acrobot to rotate freely for 15 seconds, then push it the
opposite way.
The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
responds by readily changing direction, and quickly achieves its maximum speed
with energy \(E(0,0.195)\).

>>>>

Defining \(M(q) := D_s(q)\) and \(V(q) := P_s(q)\), we find the conjugate of momenta 
is \(p = (p_u, p_a) = M(q)\dot{q}\).
The dynamics in \((q,p)\) coordinates are given by
\begin{align}\label{eqn:acrobot-hamiltonian}
    \mathcal{H}(q,p) &= \frac{1}{2}p\tpose \Minv(q) p -
    mgl\left(2 c_u + c_{ua}\right)
    , \\
     &\begin{cases}
        \dot{q} = \Minv(q) p 
        ,\\
        \dot{p}_u = -mgl\left(2s_u + s_{ua}\right) 
        ,\\
        \dot{p}_a =-\frac{1}{2}p\tpose \nabla_{q_a}\Minv(q) p
        - mgl s_{ua} + \tau,
    \end{cases} \nonumber
\end{align}
where the inverse inertia matrix is
\begin{equation}\label{eqn:Minv}
    \Minv(q) = \frac{1}{ml^2\left(2-c_a^2\right)}
    \begin{bmatrix}
        1 &
        -\left(1+c_a\right) \\
        -\left(1+c_a\right) &
        3+2c_a
    \end{bmatrix}
    .
\end{equation}
The control input is a force \(\tau \in \R\) affecting only the dynamics of
\(p_a\), representing a torque acting on the hip joint.
This means \((q,p)\) are simply actuated coordinates inside the phase space
\(\mathcal{Q} \times \mathcal{P}\) where
\(\mathcal{Q} = \mathcal{Q}_u \times \mathcal{Q}_a 
:= \Sone \times \Sone\), and
\(\mathcal{P} = \mathcal{P}_u \times \mathcal{P}_a
:= \R \times \R\).
This allows us to apply the theory of VNHCs from Chapter
\ref{ch:vnhcs}.

@@

In the spirit of perturbation analysis, we will expand our time-scaled system
\(\hat{\mu}(\alpha,\mu_0,I)\).
We begin by setting \(I = 0\) to find the nominal system
\[
\begin{cases}
    \pdiff{\hat{\mu}_0}{\alpha} = g(\alpha,\hat{\mu}_0,0)
    , \\
    \hat{\mu}_0(0) = \mu_0
   .
\end{cases}
\]
Equations ~\eqref{eqn:acrobot-alpha-dot-nom}-\eqref{eqn:acrobot-mu-dot-nom}
reveal that \(g(\alpha,\mu,0) = 0\),
so the solution to this nominal system is 
\(\hat{\mu}_0(\alpha,\mu_0) \equiv \mu_0\) for all \(\alpha\).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies \(P_\mathcal{O}(\mu_0) \leq \mu_0 - \gamma\) when \(I < 0\). 
In this case, the constraint dissipates energy on \(\mathcal{O}_1\).

@@

Indeed, let \(E_\pi\) be the level set with energy \(E(\pi,0)\), 
which forms the boundary of \(\mathcal{O}_1\). 
Define \(\bar{\mathcal{O}}_1 = \mathcal{O}_1 \cup E_\pi\) to be the closure of
\(\mathcal{O}_1\).
An orbit of the acrobot will only begin rotating if it escapes
\(\bar{\mathcal{O}}_1\) by crossing through \(E_\pi\), and eventually remains
outside this set.
It is possible for orbits starting in \(\mathcal{O}_1\) to always approach
\(E_\pi\) without ever crossing into the rotation zone.
We will prove this does not happen in general.

@@

Escaping once into the rotation domain does not guarantee the acrobot
orbit will never return to \(\mathcal{O}_1\).
Indeed, the orbit could escape through \(E_\pi\) and return to \(\mathcal{O}_1\)
several times.
However, almost every orbit must eventually remain outside the closure of 
\(\mathcal{O}_1\). 
If not, the positive limit set would once again be
\(E_\pi\); by the same argument as above, that orbit would be in \(\Pi^+\).

@@

Note that the boundary of \(\mathcal{R}\) intersects the \(p_u\) axis when 
\(p_u^2 = 60m^2 g l^3\). 
This boundary is the homoclinic orbit for the
upright equilibrium of the pendulum, which is the level set with energy
\(E(\pi,0)\).
Hence, we must have \(\rho > \sqrt{60 m^2 g l^3}\), if a
rotation has momentum \(p_u > 0\), 
and \(\rho < -\sqrt{60 m^2 g l^3}\) if it has momentum \(p_u < 0\).
The energy level set associated with \(\rho\) is 
\[
    \left\{(q_u,p_u) \in \SxR \mid E(q_u,p_u) = \frac{\rho^2}{10ml^2}\right\}
    ,
\]
which gives the relationship
\begin{equation}\label{eqn:rotation-pu2}
    \frac{p_u^2}{10m l^2} + 3mgl(1 - c_u) = \frac{\rho^2}{10 ml^2}
    .
\end{equation}
On this level set, \(q_u\) takes all values on \(\Sone\), so our angle of
rotation is uniquely parameterized by \(\beta = q_u\).
Since \(\rho\) does not change sign along the rotation, we have the smooth
relationship
\begin{align}\label{eqn:rotation-Tinv}
    q_u &= \beta
    , \\
    p_u &= \sign{\rho}\sqrt{\rho^2 - 30 m^2 g l^3\left(1 - c_\beta \right)}
    .
\end{align}

@@

Computing the acrobot's constrained dynamics in \((\beta,\rho)\)-coordinates and
setting \(I = 0\) yields the dynamics of the nominal pendulum.
MATLAB evaluates those dynamics as
\begin{align}
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\beta} &=  \sign{\rho} 
    \frac{\sqrt{\rho^2 - 30m^2gl^3(1 - c_\beta)}}{5ml^2}
    , \\
    \label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\rho} &= 0
    .
\end{align}
As expected, \(\dot{\beta}\) does not change sign on \(\mathcal{R}\) because the
orbits always flow clockwise.
If \(\rho > 0\), the rotation curve goes from \(\beta = -\pi\) to 
\(\beta = \pi\); 
if \(\rho < 0\), it goes from \(\beta = \pi\) to \(\beta = -\pi\).
By continuity of ~\eqref{eqn:rotation-T}, there exists \(I_1 > 0\) small enough
that \(\dot{\beta}\) is non-zero and does not change sign on \(\mathcal{R}\) for
all \(I \in [-I_1,I_1]\).

@@

Let us pause here for a moment to remember what we are trying to accomplish.
The second part of Theorem \ref{thm:acrobot-energy-stabilization} states that,
under suitable conditions on the integral of \(b(\beta,\rho_0)\), 
the acrobot will gain energy on 
\[
    \mathcal{O}_2(\bar{\rho}) := \left\{ (q_u,p_u) \in \SxR
    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,
\]
for some \(\bar{\rho} > \sqrt{60m^2gl^3}\).
Recall from Section \ref{sec:acrobot-proof-o1} that any acrobot initialized in
\(\mathcal{O}_1\) will eventually remain in the rotation region \(\mathcal{R}\).
What remains, then, is to prove that the acrobot's orbits will also gain
energy on
\[
    \mathcal{R}_{\bar{\rho}} := \mathcal{R} \cap \mathcal{O}_2(\bar{\rho})
    .
\]
This can be proven by analyzing the Poincar\'{e} map \(P_\mathcal{R}(\rho_0)\).
Looking at Figure \ref{fig:acrobot-rhobar-regions}, it is clear that
\(P_\mathcal{R}\) has the domain
\[
    D := \left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[ \, 
    \cup \, 
    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
which is the intersection of \(\mathcal{R}_{\bar{\rho}}\) with the \(p_u\)-axis.
We label the part of \(D\) contained in the negative \(p_u\)-axis as
\[
    D^- :=\left[-\bar{\rho},-\sqrt{60m^2gl^3}\right[
    ,
\]
and the part contained in the positive \(p_u\)-axis as
\[
    D^+ := \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    ,
\]
so that \(D = D^- \cup D^+\).
If \(P_\mathcal{R}(\rho_0)\) is always further from the origin than 
\(\rho_0\) (\ie~ it is ``expanding" on \(D\)), the acrobot will be gaining energy
on \(\mathcal{O}_2(\bar{\rho})\).

@@

\subsubsection*{Energy Gain on \(\mathcal{O}_2(\bar{\rho})\)}
The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
\(\hat{\rho}(\beta,\rho_0,I)\) by studying the evolution of the discrete time
ODE
\begin{equation}\label{eqn:rhohat-discrete}
    \rho_{n+1} := P_\mathcal{R}(\rho_n) 
    = \rho_n + I\sign{\rho_n}S(\rho_n) + 
        R(\sign{\rho_n}2\pi,\rho_n,I)
    .
\end{equation}
Here, \(\rho_n\) represents the distance along the
\(p_u\)-axis when an orbit of the constrained dynamics intersects the
\(p_u\)-axis for the \(n^\text{th}\) time, assuming the orbit was initialized at
\((q_u,p_u) = (0,\rho_0)\).

@@

We begin with \(\rho_0 \in D^+\).
Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
its parameters.
Therefore, it is bounded below on the compact set
\(\rho_0 \in \left[\sqrt{60m^2gl^3},\bar{\rho}\right]\) by some value
\(\underbar{R}(I)\).
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_2 > 0\) and \(r > 0\) so that, for all \(I \in [-I_2,I_2]\),
\[
    R(2\pi,\rho_0,I) \geq \underbar{R}(I) > -I^2r
    .
\]
Since \(S(\rho_0) \geq \epsilon\), we find that
\[
    P_\mathcal{R}(\rho_0) > \rho_0 + I\epsilon -I^2r
    ,
\]
for all \(\rho_0 \in D^+\).
Choosing \(I_3 \in \,]0,I_2]\) so that
\[
    I_3\epsilon - (I_3)^2 r \geq \gamma
    ,
\] 
makes \(P_\mathcal{R}(\rho_0) \geq \rho_0 + \gamma\) for all 
\(\rho_0 \in D^+\).

Now take \(\rho_0 \in D^-\), where we wish to show that 
\(P_\mathcal{R}(\rho_0) \leq \rho_0 - \gamma\).
The remainder \(R(\beta,\rho_0,I)\) is bounded above on the compact set 
\(\rho_0 \in \left[-\bar{\rho},-\sqrt{60m^2gl^3}\right]\), and 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_4 > 0\) and \(r^\prime > 0\) so that, for all \(I \in [-I_4,I_4]\), 
\[
    R(-2\pi,\rho_0,I) < I^2 r^\prime
    .
\]
Hence,
\begin{align*}
    P_\mathcal{R}(\rho_0) < \rho_0 - I\epsilon + I^2r^\prime
    .
\end{align*}
Choosing \(I_5 \in ]0,I_4]\) so that 
\[
    I_5 \epsilon - (I_5)^2r^\prime \geq \gamma
    ,
\] 
makes \(P_\mathcal{R}(\rho_0) \leq \rho_0 - \gamma\) for all 
\(\rho_0 \in D^-\).

@@

By the same arguments presented in this section, the Poincar\'{e} section
satisfies 
\(\abs{P_\mathcal{R}(\rho_0)} \leq \abs{\rho_0} - \gamma\)
when \(I < 0\).

@@

\section{Experimental Results}
In this section, we test our VNHC on the physical acrobot built by
\citet{xingbo_thesis} (Figure \ref{fig:xingbo-acrobot}).
Since it is not a simple acrobot, this experiment will test whether VNHCs are
robust to model mismatch.
As we will see, our VNHC does inject energy on the general acrobot model.

@@

\begin{table}
    \centering
    \caption{Physical parameters for the real acrobot, as measured by
    \citet{xingbo_thesis}.}
    \label{tab:acrobot-parameters}
    \begin{tabular}{ccccccccc}
        \toprule
        $m_u$ & $m_a$ & $l_u$ & $l_a$ & $l_{c_u}$ & $l_{c_a}$ & $J_u$ & $J_a$ & $g$ \\
        \midrule
        0.2112 & 0.1979 & % mass
        0.148 & 0.145 & % total length
        0.073 & 0.083 & % COM distance
        0.00129 & 0.00075 & % moment of inertia
        9.81 \\ % gravity
        \bottomrule
    \end{tabular}
\end{table}

@@

To compute the energy of the nominal pendulum, we evaluate the system at the
VNHC \(q_a = 0\), which yields
\[
    E(q_u,p_u) \approx 396.5501 p_u^2 + 0.5997(1 - \cos(q_u))
    .
\]
Hence, the level set of \(E_\pi\) of the energy \(E(\pi,0)\) is parameterized by
\begin{align*}
    p_u &\approx \pm \sqrt{\frac{1.1994}{793.1001}(1 + \cos(q_u))}
    ,
\end{align*}
for each \(q_u \in ]-\pi,\pi]\).

@@

\subsection{Physical Experiments}
To complete this chapter, we perform four experiments with the physical
acrobot:
\begin{enumerate}
    \item An undisturbed test to validate that the acrobot reaches rotations.
    \item A test where we stop the acrobot once it has started rotating.
    \item A test where we push the acrobot in its direction of motion after it has
        started rotating.
    \item A test where we push the acrobot against its direction of motion.
\end{enumerate}

@@

\subsubsection*{Test 2: Stopping the Acrobot}
For this test we initialize the acrobot at 
\((q_u,p_u) \approx \left(\pi,0\right)\) and let it run for 15
seconds, then stop the acrobot as it reaches the bottom of its arc.
The resulting orbit is shown in Figure \ref{fig:acrobot-stopped-orbit}.
The blue curves correspond to the orbit before the disturbance, while the red
spiral shows that the acrobot begins oscillating after it is stopped.
Despite the disturbance, it gains energy and eventually starts rotating
again.

@@

\subsubsection*{Test 3: Pushing the Acrobot Forwards}
To see how the acrobot responds when pushed in its direction of motion, we allow
the acrobot to rotate undisturbed for 15 seconds and then give it a push in its
direction of motion.
The orbit in Figure \ref{fig:acrobot-fpush-orbit} shows that the acrobot speeds
up to rotate with energy \(E(0,0.22)\), but then slows down until it reaches a
stable rotation with energy \(E(0,-0.195)\).

@@

\subsubsection*{Test 4: Pushing the Acrobot Backwards}
In this final experiment, we test whether the acrobot can easily change
directions when pushed against its current direction of motion.
We again allow the acrobot to rotate undisturbed for 15 seconds, then push it
the opposite way.
The orbit in Figure \ref{fig:acrobot-rpush-orbit} demonstrates that the acrobot
responds by readily changing direction, and quickly achieves its maximum speed
with energy \(E(0,0.195)\).

### 2020-12-22 latex/conclusion.tex
Final edit, she be complete!

This thesis has shown the utility of VNHCs as a means of injecting energy into
mechanical systems.
In Chapter \ref{ch:vnhcs} we developed the framework of VNHCs for underactuated
mechanical systems, with a focus on VNHCs for simply actuated Hamiltonian
systems.
We then applied this framework to two benchmark systems: the variable-length
pendulum and the acrobot.
In Chapter \ref{ch:vlp} we proved that a certain class of VNHCs will always
inject energy into the VLP, while in Chapter \ref{ch:acrobot} we designed a
constraint inspired by gymnastics and proved it injects energy into the
acrobot.
At the end of each chapter we performed simulations and experiments which
validated the theory and showed the utility of VNHCs. 
In the end, we demonstrated that virtual nonholonomic constraints are capable of
injecting and dissipating energy in a robust manner, all while producing
realistic biological motion.

\section{Limitations and Future Research}
Our VNHC framework relies on the following assumptions:
\begin{enumerate}
    \item The input matrix \(B(q) \equiv B \in \R^{n \times k}\) is constant and
        full rank.
    \item The input matrix has a left-annihilator 
        \(B^\perp \in \R^{(n-k)\times n}\). 
    \item The annihilator matrix \(B^\perp\) is right semi-orthogonal.
    \item The inertia matrix of the system satisfies 
        \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
    \item On the constraint manifold \(\Gamma\), one can solve for \(q_a\) as a
        function of \((q_u,p_u)\).
\end{enumerate}
If any of these assumptions are not satisfied, one may not be able to find the
constrained dynamics in simply actuated coordinates, and the results of this
thesis may not hold.
In particular, the theoretical guarantees of this thesis do not apply to systems
with physical nonholonomic constraints (\eg friction), nor do they apply to vehicles
with wheels.

>>>>

This thesis has shown the utility of VNHCs as a means of injecting energy into
mechanical systems.
In Chapter \ref{ch:vnhcs} we developed the framework of VNHCs for underactuated
mechanical systems, with a focus on VNHCs for simply actuated Hamiltonian
systems.
We then applied this framework to two benchmark systems: the variable-length
pendulum and the acrobot.
In Chapter \ref{ch:vlp} we proved that a certain class of VNHCs will always
inject energy into the VLP, while in Chapter \ref{ch:acrobot} we designed a
constraint inspired by gymnastics and proved it injects energy into the
acrobot.
At the end of each chapter we performed simulations and experiments which
validated the theory, and which showed the efficacy and robustness of VNHC-driven motion. 
In the end, we demonstrated that virtual nonholonomic constraints are capable of
injecting and dissipating energy in a robust manner, all while producing
realistic biological motion.

\section{Limitations and Future Research}
Our VNHC framework relies on the following assumptions:
\begin{enumerate}
    \item The Hamiltonian of the system is of the form 
        \[
            H(q,p) = \frac{1}{2}p\tpose \Minv(q) p + V(q)
            .
        \]
    \item The input matrix \(B(q) \equiv B \in \R^{n \times k}\) is constant and
        full rank.
    \item The input matrix has a left-annihilator 
        \(B^\perp \in \R^{(n-k)\times n}\). 
    \item The annihilator matrix \(B^\perp\) is right semi-orthogonal.
    \item The inertia matrix of the system satisfies 
        \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
    \item On the constraint manifold \(\Gamma\), one can solve for \(q_a\) as a
        function of \((q_u,p_u)\).
\end{enumerate}
If any of these assumptions are not satisfied, one may not be able to find the
constrained dynamics in simply actuated coordinates, and the results of this
thesis may not hold.
In particular, the theoretical guarantees of this thesis do not apply to systems
with physical nonholonomic constraints (\eg friction), nor do they apply to vehicles
with wheels.
